Integration of national demographic-disturbance relationships and local data can improve caribou population viability projections and inform monitoring decisions

### 2025-02-21 analysis/paper/supplement.Rmd
preparing EI submission

resultScenario = "S11"
noBiasResultScenario = "S14"

@@

Table A1. Variables and parameters in the Bayesian population model and the national demographic model.

@@

@@

```{r figDisturbanceAnthroStatic, fig.cap="Static anthropogenic disturbance and monitoring scenarios.", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/distScnscmult6ay1aSf0.png")))

@@

```{r powercmult6ay1aSf0, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure \\@ref(fig:figDisturbanceAnthroStatic) for static anthropogenic disturbance scenarios). In these examples there are 6 cows per collared cow in recruitment surveys. See Figure 4 for additional details. ", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/powercmult6ay1aSf0.png")))
```

@@

```{r figDisturbanceAnthroDeclining, fig.cap="Declining anthropogenic disturbance and monitoring scenarios.", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/distScnscmult6ay1aSf-1.png")))
```

@@

```{r powercmult6ay1aSfm1, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure \\@ref(fig:figDisturbanceAnthroDeclining) for declining anthropogenic disturbance scenarios). In these examples there are 6 cows per collared cow in recruitment surveys. See Figure 4 for additional details. ", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/powercmult6ay1aSf-1.png")))
```

@@

# Sensitivity to unknown bias in recruitment estimates from calf:cow surveys

Adding unknown unknown bias does increase uncertainty, but these increases are not large (compare Figs 6 and \@ref(fig:figStatusErrorsNoBias)). A key reason for this result is that we assumed the same range of values for the probability of misclassifying adults, and the probability of missing calves (supplement B). Presuming that surveyed groups contain more cows than bulls, and that there is an equal probability of misclassifying cows and young bulls and vice versa, then misclassification errors cause high estimated recruitment. Missing calves causes low estimated recruitment, and together these errors tend to cancel one another out.

@@

```{r figStatusErrorsLowNoBias, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",noBiasResultScenario,"/powerLamcmult6ay1aSf1.png")))
```

```{r figProbViabilityNoBias, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",noBiasResultScenario,"/probViabilitycmult6ay1aSf1.png")))
```

>>>>

resultScenario = "S15"
noBiasResultScenario = "S14"

@@

Table A1. Variables and parameters in the Bayesian population model and the national demographic model. Bayesian model estimates are distinguished from observed values, estimated expected values, true values, and true expected values using $\hat{\text{ }}$ $\tilde{\text{ }}$ $\dot{\text{ }}$ and $\bar{\text{ }}$ symbols respectively; for example, $S_t$ is the Bayesian estimate of survival in year $t$, $\hat{S}_t$ is observed survival, $\tilde{S}_t$ is the Bayesian estimate of expected survival (without stochastic interannual variation), $\dot{S}_t$ is true survival, and $\bar{S_t}$ is true expected survival (without stochastic interannual variation).

@@

For this analysis we assume the same range of values for the probability of misclassifying adults, and the probability of missing calves (Figure \@ref{fig:biasDistribution}). Presuming that surveyed groups contain more cows than bulls, and that there is an equal probability of misclassifying cows and young bulls and vice versa, then misclassification errors cause high estimated recruitment. Missing calves causes low estimated recruitment, and on average these errors tend to cancel one another out, but the set of example trajectories does include cases with high and low bias terms (Figure \@ref{fig:biasDistribution}).

@@

```{r figDisturbanceAnthroStatic, fig.cap="Static anthropogenic disturbance and monitoring scenarios.", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/distScnscmult9ay1aSf0.png")))

@@

@@

```{r figDisturbanceAnthroDeclining, fig.cap="Declining anthropogenic disturbance and monitoring scenarios.", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/distScnscmult9ay1aSf-1.png")))
```

@@

@@

# Effects of unknown bias in recruitment estimates from calf:cow surveys

Adding unknown bias does increase uncertainty and reduce accuracy, but these increases are not large (compare Figs 5 to 9 with Figs \@ref(fig:figDiffsNoBias) to \@ref(fig:figProbViabilityNoBias)) given our assumptions about the the range of plausible values for the probability of misclassifying adults, and the probability of missing calves (supplement B). In both versions of the model the estimated posterior probability of viability approaches 0 or 1 accuracy increases to 100% (compare Fig 9 and \@ref(fig:figProbViabilityNoBias)). Note that these conclusions are from a case in which assumed the same range of values for the probability of misclassifying adults, and the probability of missing calves (supplement B).

@@

```{r figEVSIsNoBias, fig.cap="The expected value of sample information without bias in recruitment estimates from calf:cow surveys. In these examples there are 6 cows per collared cow in recruitment surveys, and disturbance increases over time (Figure 1). See Figure 7 for additional details.", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",noBiasResultScenario,"/EVSIcmult6ay1aSf1.png")))
```

```{r figStatusErrorsLowNoBias, fig.cap="Variation in the probability of a correct status assessment with the true growth rate of the population the low anthropogenic disturbance scenario without bias in recruitment estimates from calf:cow surveys. In these examples there are 6 cows per collared cow in recruitment surveys, and disturbance increases over time (Figure 1). See Figure 8 for additional details.", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",noBiasResultScenario,"/powerLamcmult6ay1aSf1.png")))
```

```{r figProbViabilityNoBias, fig.cap="Relationship between posterior probability that the population is viable ($P(\\psi|x,\\tau)$) and the probability of a correct status assessment (estimated from the 114,000 example cases in Figures \\@ref(fig:figDiffsNoBias) to Figures \\@ref(fig:figEVSIsNoBias) with a generalized additive model) without bias in recruitment estimates from calf:cow surveys.", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",noBiasResultScenario,"/probViabilitycmult6ay1aSf1.png")))
```
